---
title: "tidytuesday1"
author: "Dylan Wikel"
format: html
editor: visual
---

## Libraries and Data

```{r}
library(tidyverse)
library(tidytuesdayR)

raw_data <- tt_load("2026-09-01")

world_castles_raw <- raw_data
world_castles <- world_castles_raw$world_castles
```

## Viewing Data

```{r}
glimpse(world_castles)
```

## Which Countries Have the Most Castles?

```{r}
world_castles |>
  count(sort(country)) |>
  filter(n > 200) 

world_castles |>
  count(country) |>
  filter(n > 200) |>
  ggplot(
    aes(x = reorder(country, n), y = n, 
    fill = country)) +
  geom_col()
```
France has the most castles at 351. I subset the data to countries that contain 200 or more castles after viewing all of the data in with a simple bar plot. 

## Palaces vs. Fortresses

```{r}
pvf_summary <- world_castles |>
  filter(category %in% c("palace", "fortress")) |>
  group_by(category) |>
  summarise(
    mean_year = round(mean(year, na.rm = TRUE)),
    median_year = median(year, na.rm = TRUE),
    quantile(year, 0.25, na.rm = TRUE),
    quantile(year, 0.75, na.rm = TRUE)
    )

pvf_boxplot <- world_castles |>
  filter(category %in% c("palace", "fortress")) |>
ggplot(aes(x = category, y = year)) +
  geom_boxplot()

pvf_summary
pvf_boxplot
```

Fortresses had a mean construction year of `r pvf_summary$mean_year[pvf_summary$category == "fortress"]` and a 
median year of `r pvf_summary$median_year[pvf_summary$category == "fortress"]`.